### 2011-07-19 projects/slrsim/paper/slrsim.tex
Manuscript revs

\documentclass{article}
\usepackage{natbib,graphicx,dblfloatfix,multirow,array,threeparttable}
\usepackage{xspace,bm,amsmath,amssymb}
\usepackage[nolists,tablesfirst,nomarkers]{endfloat}
\usepackage[a4paper]{geometry}
\usepackage{color}

\definecolor{red}{rgb}{1,0,0}

\linespread{1.5}

\setlength{\bibsep}{0.0pt}

\begin{document}

\newcommand{\mtwo}{PAML M2a\xspace}
\newcommand{\meight}{PAML M8\xspace}
\newcommand{\pranka}{PRANK$_{\textrm{AA}}$\xspace}
\newcommand{\prankc}{PRANK$_{\textrm{C}}$\xspace}
\newcommand{\omg}{\bm{\omega}\xspace}
\newcommand{\tpr}{TPR$_{1\%}$\xspace}
\newcommand{\sw}{sitewise\xspace}
\newcommand{\Sw}{Sitewise\xspace}
\newcommand{\Dr}{{\em Drosophila}\xspace}

\title{The effects of alignment error and alignment filtering on the \sw detection of positive selection}

\author{{Gregory Jordan and Nick Goldman}\\
\affiliation{European Bioinformatics Institute, Wellcome Trust Genome Campus, Hinxton, Cambridgeshire, CB10 1SD, United Kingdom}}
%\authorlist{Jordan and Goldman}
%\shorttitle{Effect of Alignment Error on Detecting Sitewise Selection}

%\mbeabstract{% 
\section*{Abstract}

The effects of alignment error on many types of evolutionary inference
are not well understood. Alignment filters are commonly applied to try
to reduce errors resulting from misalignment, but neither the
prevalence of such errors nor the ability of filters to reduce them
has been systematically studied. Focusing on the sitewise detection of
positive selection, we performed simulation experiments to quantify
the false positives and false negatives introduced by alignment error
and the ability of alignment filters to improve performance. We found
that some aligners led to many false positive results across a wide
range of divergence levels, while others resulted in very few. False
negatives were a problem for all aligners, increasing with
divergence. Of the aligners tested, PRANK consistently performed the
best and ClustalW performed the worst. The best alignment filters
showed good ability to reduce the error rates from ClustalW
alignments, but they failed to improve the performance of PRANK
alignments under most circumstances. Of the filters tested, GUIDANCE
performed the best and Gblocks performed the worst, failing to improve
even the error-prone ClustalW alignments. The performance of the true
alignment and an optimal filtering method suggested that performance
improvements could be gained by improving aligners or filters to
reduce the prevalence of false negatives, especially at higher
divergence levels and indel rates. \textcolor{red}{As no combination
  of aligner plus filter significantly outperformed the best aligner
  alone,} we conclude that for the sitewise evolutionary analysis of
most vertebrate, insect and fungal phylogenies, the use of unfiltered
PRANK alignments is recommended.

%}

%\keywords{Alignment, positive selection, alignment filtering, indels}

%\email{greg@ebi.ac.uk}

%\mbestyle

\section*{Introduction}

The decreasing cost of DNA sequencing has triggered a striking
increase in the number of model and non-model organisms with planned
genome sequencing projects, suggesting that the range and scale of
comparative genomics applications will continue to expand
\citep{Green20072x,2007Identification}. These clusters of
closely-related genome sequences across a wide taxonomic range have
led to a better understanding of which aspects of molecular evolution
are variable and which are constant \citep{Wolf2009Nonlinear} and an
increased sampling of species should continue to boost the power and
accuracy of individual analyses within a given clade.

The study of protein evolutionary rates and selective pressures in
particular has flourished as a result of the growth in comparative
genomics datasets. This is especially beneficial for the calculation
of spatially precise evolutionary estimates, as additional species
sampling has been shown to be an effective means of boosting the
accuracy and power of \sw detection of positive selection and
evolutionary constraint
\citep{Anisimova2001Accuracy,Massingham2005Detecting}. Site-specific
evolutionary estimates have proved especially valuable when analyzed
in conjunction with other protein-based datasets such as protein
structural features \citep{Lin2007Proportion,Ramsey2011Relationship},
human population diversity \citep{2010Map} and human disease
mutations \citep{Arbiza2006Selective}.

A major concern in the detection of protein positive selection is that
the effect of alignment error is not well characterized. Intuitively,
one might expect alignment error to result mainly in an increased
number of false positives, as the spurious alignment of nonhomologous
codons on average would result in a high number of apparent
non-synonymous substitutions and a low number of synonymous
substitutions (since a randomly chosen pair of codons are more likely
to be nonsynonymous than synonymous). However, false negatives may
also be introduced, either through over-alignment of synonymous codons
(increasing the synonymous substitution rate) or non-alignment of
homologous codons (causing reduced power due to less evolutionary
information). Since different aligners employ a variety of algorithms,
evolutionary models, and heuristic optimizations
\citep{Notredame2007Recent}, each program may be more or less prone to
different types of alignment error, causing potentially large
variations in the nature and magnitude of its impact on the detection
of positive selection.

\input{table1.txt}

The protein structure and the evolutionary divergence of a dataset may
also contribute to the effects of alignment error. Differently
structured protein regions show variable tolerance to biological
indels, with indels more common in extracellular and transmembrane
proteins than in highly folded enzymes and housekeeping genes
\citep{delaChaux2007DNA} . This suggests that well-folded protein
regions will experience fewer biological indels---and will therefore
be less susceptible to alignment error---than less structured
regions.

Furthermore, the evolutionary divergence of a dataset affects the
power of \sw inference and the prevalence of indels in multiple
ways. As maximum-likelihood methods for detecting positive selection
require data in the form of observed substitution events, they show
little power at low divergence and their highest power at intermediate
to high divergence levels \citep{Anisimova2001Accuracy}. However,
alignment error should be greatest at high divergences, which may have
the effect of reducing power. These two trends suggest that the
overall power will be low at both extremes of divergence, with little
inference power at low divergence (due to the scarcity of data in the
form of observed substitutions) and an overwhelming amount of
alignment error at high divergence (due to the large number of indel
events).

Luckily the majority of genes in many biological clades of interest
(such as mammals, vertebrates, fruit flies, and yeast) fall within the
middle range of divergences where \sw methods are at their most
powerful and where multiple alignment is a difficult---but not
hopeless---problem. As such, it is important to seek an understanding
of the impact of alignment error on overall error rates within this
important range of divergence levels.

A number of empirical analyses have established that errors in gene
sequencing, annotation and alignment can contribute to errors in
downstream evolutionary analyses such as phylogeny inference
\citep{Wong2008Alignment} and estimates of positive selection
\citep{Schneider2009Estimates,Markova-Raina2011High}. Most recently,
Markova-Raina and Petrov \citeyearpar{Markova-Raina2011High} showed
that the detection of positively-selected sites and genes in \Dr
genomes is highly sensitive to aligner choice, with PRANK's codon
model \citep{Loytynoja2008PhylogenyAware} consistently producing
alignments with the lowest amount of positive selection. Still,
according to the authors' manual inspection of alignments, even
positively-selected sites identified with PRANK alignments contained a
sizable proportion of apparent false positives.

A limitation of the analysis of error in empirical datasets is the
lack of a benchmark set of true alignments and positively-selected
sites. Markova-Raina and Petrov \citeyearpar{Markova-Raina2011High}
used their expected general effect of alignment error (an increase in
false positives due to misalignment of non-homologous codons) as a
proxy by which to compare different methods, allowing for the
conclusion that PRANK was the least error-prone aligner in their
analysis. However, the absolute number of false positives remained
uncertain and there was the possibility of conflating multiple sources
of error: in addition to alignment error, the authors noted that gene
mis-annotation was responsible for many apparent false positives, and
there is also an expected error rate from the likelihood inference
method itself. This limitation leaves important and interesting
questions, regarding the nature of alignment error and its
quantitative impact on the detection of positive selection unanswered
by empirical studies.

Controlled simulation experiments provide a natural framework for
investigating error rates in detail, allowing one to pinpoint the
sources of error in multi-step analyses such as alignment followed by
evolutionary inference. This approach has been employed in assessing
the robustness of phylogenetic inference methods to misalignment
\citep{Dwivedi2009Phylogenetic,Ogden2006Multiple,Loytynoja2008PhylogenyAware}
but those results cannot be easily extrapolated to the analysis of \sw
selective pressures. More recently, Fletcher and Yang
\citeyearpar{Fletcher2010Effect} performed a series of simulation
experiments investigating alignment error in the use of the
branch-site test to detect positive selection in genes. Their results
showed that most aligners caused false positives by over-aligning
codons and that datasets from mammalian and vertebrate gene families
contain enough evolutionary divergence to make false positive errors
resulting from misalignment a legitimate concern.

Reflecting a widespread awareness of the problem of misalignment,
methods for identifying and removing uncertain or unreliable alignment
regions have been commonly used in phylogenetic and molecular
evolutionary analyses. The popular Gblocks program applies a set of
heuristic criteria to identify conserved blocks deemed suitable for
phylogenetic or evolutionary analysis \citep{Castresana2000Selection}
while a number of aligners such as T-Coffee
\citep{Notredame2000TCoffee} and PRANK
\citep{Loytynoja2008PhylogenyAware} produce estimates of alignment
confidence or reliability. \textcolor{red}{GUIDANCE, which measures
  the robustness of alignment regions to perturbations in the guide
  tree used for progressive alignment, has also been proposed as an
  alignment confidence score
  \citep{Penn2010Alignment}}. Unfortunately, despite their widespread
use, the impact of the many available \textcolor{red}{alignment
  scoring and filtering} methods on phylogenetic and evolutionary
analyses has not been well studied. Even for a single filtering
program, Gblocks, results have been contradictory: one
simulation-based study found that it improved the phylogenetic signal
\citep{Talavera2007Improvement} while an empirical study across a wide
range of taxa found that Gblocks-filtered alignments produced worse
phylogenetic trees than unfiltered alignments
\citep{Dessimoz2010Phylogenetic}. With the application of published
filtering methods to alignments before testing for positive selection
becoming standard practice
\citep{Studer2008Pervasive,Aguileta2009Rapidly}, an analysis of the
benefits of alignment filtering to the detection of positive selection
would seem well-warranted.

This paper aims to use a simulation framework to incorporate alignment
error and alignment filtering into estimates of the error rate and
power of \sw evolutionary inference of positive selection. Our
approach is similar to that of Anisimova et al.
\citeyearpar{Anisimova2002Accuracy} in that we use simulated protein
alignments to evaluate methods for detecting \sw positive selection,
considering each site to be an independent hypothesis test producing a
positive or negative result. Our experimental design also shares some
features with Fletcher and Yang's \citeyearpar{Fletcher2010Effect}
investigation of the performance of branch-site method for detecting
genewise positive selection under alignment error. Like Fletcher and
Yang \citeyearpar{Fletcher2010Effect}, we simulate codon alignments
with indels and calculate inferred alignments using a small but
diverse sample of aligners. Unlike their study, here we focus on \sw
detection of positive selection occurring throughout a phylogeny (as
opposed to genewise detection of selection acting at specific
branches), we explore a wider range of plausible tree sizes, indel
rates and divergence levels, and we evaluate the impact of a number of
alignment filtering methods on the \sw analysis. Given the recent
proliferation of completed and planned vertebrate genomes, we pay
special attention to choosing simulation parameters and analysis
methods similar to those commonly encountered in the \sw analysis of
vertebrate gene families.

\section*{Methods}

\input{table2.txt}

\subsection*{Alignment Simulations}

An overview of the simulation parameters used in this study can be
found in Table \ref{table_1}. Three rooted trees were used to guide
the simulation of protein-coding DNA alignments: the artificial
6-taxon tree used by Anisimova et al.
\citeyearpar{Anisimova2001Accuracy} and Massingham and Goldman
\citeyearpar{Massingham2005Detecting} rooted at its midpoint, the
17-taxon vertebrate $\beta$-globin tree from Yang et
al. \citeyearpar{Yang2000CodonSubstitution} and the 44-taxon
vertebrate tree used by the ENCODE project
\citep{2007Identification,Nikolaev2007Early}. Trees, shown with their
original branch lengths in Figure \ref{fig_1}, were scaled to
comparable divergence levels by normalizing their mean path length
(MPL), defined as the root-to-tip branch length averaged across all
lineages in the tree. We simulated alignments with MPL divergence
between 0.05 and 2.0 synonymous substitutions per synonymous site,
spanning the range of evolutionary divergences observed in several
clades of organisms with fully-sequenced genomes (Table
\ref{table_2}).
\begin{figure}[t]
\begin{center}
%\includegraphics[scale=0.5]{fig1.pdf}
\includegraphics[scale=0.8]{fig1.pdf}
\end{center}
\caption{Phylogenetic trees used for simulation and analysis. The
  original scale for each tree is indicated by a scale bar, but trees
  were scaled to equal mean path length (MPL) divergence levels for
  simulation. (a) A 6-taxon artificial tree used in previous
  simulations
  \citep{Anisimova2001Accuracy,Massingham2005Detecting}. (b) A tree
  estimated from $\beta$-globin genes of 17 vertebrates and used in
  previous empirical analyses and simulation studies
  \citep{Anisimova2001Accuracy,Anisimova2002Accuracy}. (c) The
  44-species tree used by the ENCODE project
  \citep{2007Identification,Nikolaev2007Early}. The nodes indicated by
  arrows were used as the reference species when comparing the true
  and inferred alignment (see Methods).}
\label{fig_1}
\end{figure}

The Indelible program \citep{Fletcher2009INDELible} was used to
simulate codon sequences with indels along each phylogenetic tree. The
length of the root sequence was set to 500 codons and $\bm{\kappa}$ (the
ratio of transition to transversion substitutions) was fixed at 4.
Indel lengths were drawn from a discretized
power-law distribution with an exponential decay parameter of 1.8 and
a maximum value of 40, yielding a mean indel length of 3.33 codons and
standard deviation of 5.51 codons. This power-law model of indel
lengths is well-supported by empirical studies
\citep{Benner1993Empirical,Cartwright2009Problems} and manual
inspection of alignments from a range of parameter values identified
the chosen model parameters as resulting in alignments most closely
resembling those encountered in vertebrate alignments. The ratio of
insertion to deletion events was set to 1, and the rate of indel
formation was varied between 0 and 0.2 indel events per substitution
per site.

The distribution of \sw selective pressures (embodied by the parameter
$\omg$, the ratio of the rate of non-synonymous substitution to the
rate of synonymous substitution) was modeled with a log-normal
distribution derived from a maximum-likelihood fit to a large dataset
of \sw selective pressures estimated from mammalian gene trees
(\citeyearpar{LindbladToh2011High}; log-normal parameters shown in Table
\ref{table_1}). This distribution, with mean $\omg$ of 0.28 and 6\% of
sites having $\omg>1$, is consistent with the structure-based
expectation of many protein sites under purifying selection and few
under neutral selection or positive selection
\citep{Smith1970Natural,Kimura1974SomePrinciples}. Indelible's general
discrete model of \sw $\omg$ variation was used to approximate the
log-normal distribution by splitting the probability density into 50
equally-spaced bins between $\omg$ values of 0 and 3, with the highest
bin containing the probability density for all values $\omg>3$.

Branch lengths for each of the simulation trees were scaled before
simulation to correct for the difference between our definition of
branch lengths as the number of synonymous substitutions per
synonymous site ($dS$) and Indelible's interpretation of branch length
as the average number of substitutions per codon ($t$)
\citep{Fletcher2010Effect}. They are related approximately by
$t=3(NdN+SdS) = 3dS(\bar{\omg{}}N+S)$, where $N$ and $S$ are the
proportion of nonsynonymous and synonymous sites and $\bar{\omg}$ is
the mean $\omg$ across all sites. $S$ is approximately 0.3 when
$\bm{\kappa}=4$ \citep{Yang1998Synonymous} and the mean $\omg$ ratio
for our chosen distribution is 0.277, yielding a $dS$-to-$t$
conversion factor of 1.48 for all simulations performed.

\subsection*{Sequence Alignment and Filtering}

Alignments were inferred using four alignment algorithms chosen for
their widespread use or demonstrated accuracy: ClustalW v1.82
\citep{Thompson1994ClustalW}, MAFFT \citep{Katoh2005MAFFT} and two
variants of PRANK \citep{Loytynoja2005From} based on an amino acid
model (subsequently referred to as \pranka{}) or an empirical codon
model (subsequently referred to as \prankc{}). Unaligned amino acid
sequences were given as input to all alignment programs (except
\prankc{}, which was provided the unaligned DNA sequences) and all
software was run using default parameters with the true phylogenetic
tree given as input where possible.

Alignments were filtered by masking out residues based on the output
of three alignment scoring methods: Gblocks conserved blocks
\citep{Castresana2000Selection}, T-Coffee consistency scores
\citep{Notredame2000TCoffee,Notredame2003Using}, and GUIDANCE
alignment confidence scores \citep{Penn2010Alignment}. Gblocks, which
identifies entire alignment columns as conserved or not conserved, was
run using an increased gap tolerance and a reduced minimum block
length in order to reduce the amount of each alignment removed
(command-line parameters {\em b5=a} and {\em b4=3}), and all residues
from any columns not within an identified conserved block were masked
with $N$s.

\textcolor{red}{In contrast to Gblocks, GUIDANCE and T-Coffee produce
  individual scores for each residue,} \textcolor{red}{ allowing for
  individual residues to be masked instead of entire columns. GUIDANCE
  generates many replicate alignments, each using a slightly perturbed
  guide tree, and assigns to each residue a score from 0 to 1 based on
  how consistently it was placed in the replicate alignments. We ran
  GUIDANCE with 100 replicates using MAFFT as the aligner, except when
  \prankc was the original aligner, in which case 30 replicate
  alignments were produced using \pranka. T-Coffee calculates the
  consistency between an input multiple alignment and independently
  calculated pairwise alignments \citep{Notredame2003Using}, rounding
  and normalizing residue scores into integers between 0 and
  9. T-Coffee was run using its default settings and the {\em
    evaluate\_mode --output=score\_ascii} command-line parameters to
  output alignment scores.}

\textcolor{red}{To filter alignments based on these residue-wise
  scores, a cutoff threshold was chosen for each method (0.5 for
  GUIDANCE and 5 for T-Coffee) and residues equal to or below that
  threshold were masked. On a per-alignment basis, if the default
  threshold caused greater than 50\% of residues to be masked, then
  the threshold was relaxed to the highest value for which at least
  50\% of residues remained unmasked. We found this adjustment
  necessary because the scores from GUIDANCE and T-Coffee were
  strongly affected by the simulation conditions, with much lower
  average scores at higher indel rates and divergences. Requiring at
  least 50\% of residues to remain unmasked ensured that enough data
  were available for meaningful evolutionary analysis, mimicking
  typical treatment of real data sets.}

Two unrealistic but informative datasets were produced to serve as controls. First,
the true simulated alignment was included in order to evaluate the \sw
performance without any alignment error. Second, an additional
filtering method was constructed to represent an unattainable
best-case scenario for sequence filtering, using knowledge of the true
alignment to assign a score to each residue reflecting how correctly
is has been placed in the inferred alignment. The approach taken was
to calculate, for each residue, the branch length of the correct
sub-tree (defined as the sub-tree connecting all sequences to which
the current residue was correctly aligned) divided by the branch
length of the total aligned sub-tree (defined as the sub-tree
connecting all sequences with non-gap residues at the current
alignment column). This residue-wise score ranges from 0 to 1 and
reflects the expectation that correctly-aligned evolutionary branch
length is the main source of information from which \sw inference
methods derive their power. We refer to this method as the 'optimal'
filtering method. Scores were handled in a manner similar to GUIDANCE and
T-Coffee, using a score threshold of 0.5.

\subsection*{\Sw Evolutionary Analysis}

\Sw estimates of selective pressures were calculated using
maximum-likelihood methods implemented in the Phylogenetic Analysis by
Maximum Likelihood (PAML; \citealt{Yang2007PAML}) and Sitewise Likelihood Ratio
(SLR; \citealt{Massingham2005Detecting}) software packages.

The $codeml$ program from PAML implements a number of likelihood ratio
tests (LRTs) for detecting the presence of positive selection in a
gene while allowing the $\omg$ ratio to vary among sites
\citep{Yang2000CodonSubstitution}. These models, known as the sites or
random sites models, use a variety of predefined statistical
distributions to account for heterogeneous $\omg$ ratios among sites. After
the likelihood optimization is performed, Bayesian methods can be used
to estimate the posterior probability of each site being drawn from a
given site class, where a high posterior probability of a site
belonging to a class with $\omg>1$ can be considered strong
evidence that a site has evolved under positive selection
\citep{Yang2005Bayes}. We used the two models for which the
recommended Bayes Empirical Bayes method are implemented, M2a and M8.

SLR implements a method specifically designed for \sw estimates which
has been shown in simulations to perform as well as or better than
PAML's \sw random sites models \citep{Massingham2005Detecting}. SLR
models codon evolution as a continuous-time Markov process where
substitutions at one site are independent of substitutions at all
other sites. No assumptions are made regarding the
distribution of $\omg$ ratios within the alignment. The value of
$\omg$ is considered to be an independent parameter at each site:
after first optimizing shared parameters using the whole alignment,
SLR uses the shared parameters and the data at each alignment site to
calculate a \sw statistic for non-neutral evolution. This statistic is
based on a likelihood-ratio test where the null model is neutral
evolution ($\omg=1$) and the alternative model is either purifying or
positive selection ($\omg<1$ or $\omg>1$, respectively). The raw
statistic measures the strength of evidence for non-neutral evolution
at each site; following Massingham and Goldman
\citeyearpar{Massingham2005Detecting} we use a signed version of the
SLR statistic (created by negating the statistic for sites with
$\omg<1$) as the test statistic for positive selection.

\subsection*{Measuring Performance}

In order to compare \sw estimates from different alignments, a single
sequence from each tree was chosen as the reference (arrows, Figure
\ref{fig_1}) and all \sw statistics were mapped from alignment columns
to sequence positions in the reference sequence. This approach bears
some resemblance to the process of mapping alignment-based
evolutionary estimates onto a single member of the alignment for
further analysis and integration with other genome-referenced data (as
is often done, for example, using mammalian alignments and a human
reference). As a result of this reference sequence based mapping,
sites which were deleted in the reference sequence or inserted in a
lineage not ancestral to the reference were \textcolor{red}{not
  included in the final performance analysis.}

To evaluate the power and error rates that might be achieved in
real-world data analysis, the recommended cutoff thresholds for PAML's
Bayesian posterior probabilities and the SLR statistic were used to
identify positively selected sites. A posterior probability threshold
of 0.95 was used for PAML \citep{Yang2005Bayes} and a threshold of 3.84,
the 95\% critical value of the $\chi^2$ distribution with 1 degree
of freedom, was used for SLR \citep{Massingham2005Detecting}. Sites were
compared to their true simulated state (e.g. positively-selected or
non-positively selected) in order to identify correct and incorrect
inferences, and from these classifications we calculated the false
positive rate (FPR, defined as the proportion of all sites with true
$\omg<1$ falsely identified as positively selected) and true positive rate
(TPR, defined as the proportion of all sites with true $\omg>1$ correctly
identified as positively selected).

As the addition of alignment error is expected to affect the power and
error rates differently for each combination of simulation condition
and aligner, we identified the score thresholds for each dataset that
resulted in an actual FPR of 1\% and calculated the TPR achieved at
this actual error rate (hereafter referred to as \tpr{} to distinguish
it from the TPR described above). Although this estimate of
error-controlled power would be impossible to calculate in an
empirical analysis where the error rate is unknown, it is useful in a
simulation context for allowing a controlled comparison of the
performance of \sw analysis between different
conditions. Specifically, it should be sensitive to changes in the
numbers of both false positives and false negatives resulting from
alignment error or alignment filtering; in both cases a lowered
error-controlled power would result, as fewer true positives are
identified at the constant 1\% FPR.

We also evaluated the ability of each method to accurately infer the
$\omg$ value at each site by collecting sitewise $\omg$ estimates from
the output of each method and calculating the Pearson's correlation
coefficient between the true and inferred $\omg$ values for each set
of simulation conditions.

\begin{figure}[t]
\begin{center}
%\includegraphics[scale=0.75]{fig2.pdf}
\includegraphics[scale=1.25]{fig2.pdf}
\end{center}
\caption{Alignments were simulated without indels for three tree
  shapes and analyzed with SLR, \meight, or \mtwo. Fifty replicate
  alignments were simulated for each data point. The performance of
  each analysis method, as measured by four summary statistics, is
  plotted as a function of the mean path length (MPL) divergence. From
  top to bottom: true positive rate (TPR) at the recommended cutoff
  threshold (0.95 for PAML and 3.84 for SLR); false positive rate
  (FPR) at the recommended cutoff threshold; TPR at a 1\% FPR
  threshold; Pearson's correlation coefficient between the true and
  inferred sitewise $\omg$.}
\label{fig_2}
\end{figure}

\section*{Results and Discussion}
\subsection*{The Performance of Three Methods for Detecting \Sw Positive Selection}
We first evaluated the ability of three \sw methods, \mtwo,
\meight and SLR, to accurately estimate \sw $\omg$ values and to
detect positive selection under a range of tree lengths in the absence
of alignment error. Figure \ref{fig_2} shows the TPR, FPR, \tpr{} and \sw $\omg$
correlation over a range of MPLs for each of the three simulation
trees.

The detection power and $\omg$ correlation were weakest at low
divergence levels for all methods and all trees due to the low amount
of evolutionary information, as observed in previous simulations
\citep{Anisimova2002Accuracy}. We found a positive correlation between
tree size and detection power, with the highest performance in the
44-taxon tree. Power generally increased monotonically with
divergence, except for the 6-taxon tree which saw its maximum
performance at moderate divergence levels (MPL 0.5--1.0) and began
decreasing at higher values. The downward trend in the 6-taxon tree
was likely due to the impact of saturation of synonymous sites in the
very long branches present in such a sparse tree at high divergence
levels. With lower average branch lengths at equivalent MPLs, the two
larger trees showed no signs of decreased performance even at a MPL of
2 substitutions per site, which is greater than any of the divergence
levels found in groups of commonly analyzed vertebrate, insect and
fungal species (Table \ref{table_2}).

Comparing the three methods for detecting positively selected sites, we found that at the
recommended cutoff threshold (Figure \ref{fig_2}, top row) SLR showed
the highest power to detect positive selection in all trees, followed
by \meight and \mtwo. In the smallest tree, the power of the two
PAML methods was virtually zero while SLR reached a maximum TPR of ~18\%
(at MPL=0.5). At the same divergence, SLR yielded TPRs of 25\% and 45\%
in the 17-taxon and 44-taxon trees, respectively, with \mtwo ranging
between 50--75\% of SLR's power and \meight falling between the two
other methods.

The TPR measurements represent the power that might be achieved in
real-world analysis using recommended cutoff thresholds, but the
higher power from SLR may merely reflect a shifted balance between
power and accuracy at the recommended cutoff threshold as opposed to
an increased absolute ability to discriminate positive from neutral or
purifying selection. The FPR and error-controlled \tpr results
revealed that this was indeed the case: the FPR from SLR was higher
than that from either of the PAML methods for all trees and divergence
levels, suggesting that its higher power was the result of a
less-conservative cutoff value. This was further verified by
evaluating the TPR at a cutoff threshold that controlled for an actual
FPR of 1\% for each method (\tpr{}, third row in Figure
\ref{fig_2}). The error-controlled \tpr values were virtually
identical for all three methods, providing strong evidence that the
three methods' \sw statistics were nearly equally sensitive to
positive selection under our chosen simulation conditions.

The conservative nature of the default thresholds for PAML and SLR
has been previously noted
\citep{Anisimova2002Accuracy,Yang2005Bayes,Massingham2005Detecting},
but the extremely low false positive rates in our simulations showed
that in the absence of alignment error all three methods would yield
very few false positives when analyzing genes with a typical
mammalian-like distribution of $\omg$ values. The low FPRs were likely
due to the large proportion of sites under moderately strong purifying
selection in our $\omg$ distribution used for simulation. Such sites
are less likely to yield false positives than sites under neutral
evolution ($\omg$=1), the null model against which tests for positive
selection are traditionally controlled.

For the purposes of our indel experiments, the observed similarity in
error-controlled power levels indicated that the behavior of \mtwo,
\meight, and SLR was similar enough not to warrant separately
evaluating all three methods in the subsequent indel simulation
experiments. As the runtime for SLR was significantly lower than that
of either PAML model, all subsequent results are presented only based
on the SLR test.

\subsection*{The Effect of Alignment Error on \Sw Power}

When the indel rate was greater than zero, performance levels varied
significantly for different tree sizes, alignment algorithms, and
evolutionary divergences. Figure \ref{fig_3} shows the same
performance measurements as Figure \ref{fig_2} for simulations without
indels (gray lines, Figure \ref{fig_3}) and with indels (black lines,
Figure \ref{fig_3}) analyzed using three different aligners and the
true alignment. For the indel simulations, the indel rate here was held
constant at 0.1 indel event per substitution.
\begin{figure}[t]
\begin{center}
%\includegraphics[scale=0.75]{fig3.pdf}
\includegraphics[scale=1.25]{fig3.pdf}
\end{center}
\caption{Sequences were simulated without indels (gray lines) or with
  indels (black lines) using one of three tree shapes, aligned with
  one of three aligners, and analyzed with SLR; true alignments were
  separately analyzed with SLR (solid black lines). One hundred
  replicate alignments were simulated for each data point. The
  performance of each dataset, as measured by four summary statistics,
  is plotted as a function of the mean path length (MPL)
  divergence. From top to bottom: true positive rate (TPR) at the
  recommended cutoff threshold; false positive rate (FPR) at the
  recommended cutoff threshold; TPR at a 1\% FPR threshold; Pearson's
  correlation coefficient between the true and inferred $\omg$.}
\label{fig_3}
\end{figure}

Comparing the results without indels to those with indels under the
true alignment we found a slight decrease in power and $\omg$
correlation and no noticeable increase in FPR. The decreased power was
expected, since even in the absence of alignment error alignment
columns containing gaps harbor less evolutionary information than
columns with complete sequence data. The lack of increased FPR showed
that SLR retained its conservative statistical performance even when
analyzing gapped alignments. Surprisingly, at higher divergences (MPL
$>$ 1.0) under the six-taxon tree, the FPR with indels was lower than
the FPR without indels. This unexpected result may be attributed to
the large number of alignment columns under such conditions that
contained only a single non-gap sequence, as those columns were never
inferred as positively-selected by SLR due to the complete lack of
information. The two larger trees did not show a similar trend at high
divergence levels, suggesting that this effect was indeed due to the
highly sparse nature of the alignments in the 6-taxon tree. All other
results from the 6-taxon tree at high divergences were similarly
anomalous in this respect; we surmised that the sparseness of the true
alignment, combined with the extreme difficulty of accurately aligning
sequences along very long branches, made \sw analysis with indels very
unreliable at high divergences in the smallest tree.

When alignments were inferred using one of the three aligners tested,
the TPR, \tpr{} and $\omg$ correlation were all reduced relative to
the true alignment (dashed and dotted lines, Figure \ref{fig_3}). The
degree of reduction varied depending on the aligner, simulation
conditions, and performance measurement being analyzed. At low
divergences (e.g. MPL $<$ 0.2) the inferred alignments generally
showed only a small decrease in performance. As divergence levels
increased, so did the difference between the performance of the true
alignment and the inferred alignments. The three aligners tested could
be consistently and unambiguously ranked by all of the measured
performance characteristics, with \prankc always performing best and
ClustalW performing worst. The same ranking of aligners with respect
to detecting positive selection was found by Fletcher and Yang
\citeyearpar{Fletcher2010Effect}; our results corroborate their
finding and provide evidence that this ranking may be applicable to a
wider array of evolutionary analyses.

Looking at the TPR results for inferred alignments, we observed that
in the 6-taxon tree the three aligners formed a cluster of lines well
below the true alignment value, indicating similar tendencies among
the different aligners to produce false negatives in the smaller
tree. In larger trees the different aligners showed a wider spread of
TPR values, but even \prankc{} showed a 5--10\% reduction compared to
the true alignment at MPL$=$1.0. These results show that the
introduction of false negatives is a significant and seemingly
unavoidable result of alignment error at medium to high divergence
levels (MPL $>$ 0.5), with even the most successful aligner producing
a marked reduction in TPR compared to the true
alignment. \textcolor{red}{The \tpr and $\omg$ results in the larger
  two trees were qualitatively similar to the TPR results, showing
  that the aligners tested led to different levels of sitewise
  performance even when controlling for actual error rates or
  assessing the sitewise $\omg$ correlation.}

The FPRs for infered alignments exhibited a very different trend from
the other performance measures, with generally higher FPRs than the
true alignment and the widest range of values occurring in the 6-taxon
tree. In this tree at medium divergence levels (e.g. MPL 0.2--0.6)
ClustalW showed up to a fourfold increase, and \prankc{} a nearly
twofold increase, in FPR over the true alignment. As previously noted,
the 6-taxon tree showed an anomalous FPR pattern at higher
divergences, with lower FPRs for inferred alignments than the true
alignment, likely due to the highly sparse true alignment under those
conditions. In the two larger trees, FPRs from inferred alignments
were less elevated compared to the true alignment, less variable
between aligners, and relatively constant across the range of
divergences. ClustalW's FPR ranged between 0.001 to 0.005, while
\prankc{}'s FPR was virtually identical to that of the true alignment in the
17-taxon and 44-taxon trees.

We found it useful to combine divergence estimates from Table
\ref{table_2} with the results from Figure \ref{fig_3} to characterize
the combined effects of alignment error at different commonly analyzed
levels of divergence. For example, at a human-mouse divergence level
(MPL$=$0.2) misalignment had little impact on the TPR regardless of
what aligner was used. However, ClustalW yielded a notably higher FPR
than MAFFT or \prankc, and the error-controlled \tpr{} was
correspondingly lower for ClustalW in all three trees. Thus, at low
divergences we found that false positives were the main source of
error from misalignment, and different aligners had highly variable
tendencies to produce false positive results. At higher vertebrate and
\Dr divergence levels (MPL$=$0.8--1.0) false negatives became much
more prevalent. The TPR for all inferred alignments was virtually zero
in the 6-taxon tree, underscoring the necessity of including many
species in the analysis of highly diverged sequences. In the two
larger trees, \prankc resulted in very few additional false positives,
but it suffered a 5--10\% reduction in TPR relative to the true
alignment. Meanwhile, ClustalW showed a 50\% TPR reduction and
maintained a strongly elevated FPR. At higher divergences and in
larger trees, false negatives were thus the most persistent effect of
alignment error, causing a marked reduction in \sw power even with the
best-performing aligner. \textcolor{red}{Overall, the ranking of
  aligners was clear, with \prankc performing better than MAFFT, and
  MAFFT performing better than ClustalW, across all performance
  measures and MPL divergence levels.}

\begin{figure*}[t]
\begin{center}
%\includegraphics[scale=0.8]{fig4.pdf}
\includegraphics[scale=0.8]{fig4.pdf}
\end{center}
\caption{Sequences were simulated with indels using one of three tree
  shapes and a range of indel rates and mean path length (MPL)
  divergence levels. Alignments were inferred with one of four
  aligners and analyzed with SLR; true alignments were separately
  analyzed with SLR. One hundred replicates were simulated for each
  set of conditions. Each square is colored according to the
  performance at a given (indel rate, MPL) pair as measured by one of
  three summary statistics: (a) the true positive rate (TPR) at the
  recommended cutoff threshold, (b) the false positive rate (FPR) at
  the recommended cutoff threshold, or (c) the TPR at a 1\% FPR
  threshold.}
\label{fig_4}
\end{figure*}

\subsection*{\Sw Power Under a Range of Indel Rates and Divergences}

To explore the effects of alignment error across a wider range of
simulation conditions, we extended the simulations of Figure
\ref{fig_3} across multiple indel rates. Figure \ref{fig_4} shows
heatmaps of the TPR and FPR for ClustalW, \prankc and the true
alignment (Figure \ref{fig_4}A,B) and a heatmap of the
error-controlled \tpr for all aligners tested (Figure
\ref{fig_4}C). The results from Figure \ref{fig_3}, which were
simulated with an indel rate of 0.1, correspond to the middle row of
each panel in Figure \ref{fig_4}; rows above and below the middle row
represent higher and lower indel rates, respectively. \textcolor{red}{Similarly, the
bottom row of each panel in Figure \ref{fig_4} was simulated with an
indel rate of 0 and corresponds to the 'No Indels' data in Figure
\ref{fig_3}.}

The TPR values (Figure \ref{fig_4}A) show a consistent pattern
across the range of indel rates, with power decreasing as either the
indel rate or the divergence level increases (except at the lowest
divergence levels, where the lack of evolutionary information yielded
slightly lower TPRs in the larger two trees). \prankc{} showed a
greater ability than ClustalW to maintain a high TPR at higher indel
rates, especially in the 17-taxon and 44-taxon trees. At lower indel
rates, the TPR performance of both aligners and the true alignment
were qualitatively similar.

\prankc{} and ClustalW both showed a qualitatively similar pattern of
elevated FPRs in the 6-taxon tree (Figure \ref{fig_4}B), but their
behavior diverged significantly in the 17-taxon and 44-taxon trees. In
the 17-taxon tree, \prankc{} only showed an elevated FPR compared to
the true alignment at very high indel rates and divergence levels, but
the ClustalW FPR increased steadily with the indel rate, quadrupling
in value from the lowest to highest indel rate. Interestingly, for any
given indel rate, the ClustalW FPR showed little variation across the
range of divergence levels. This result was counter-intuitive, as we
expected alignment errors to become more common as divergence
increased and the number of observed indel events grew. Furthermore,
\prankc{} behaved according to our expectations, showing increased
FPRs only at the highest divergences and indel rates in the 17-taxon
tree. The FPR results in the 44-taxon tree confirmed the strange
effect of ClustalW’s alignments on the \sw FPR: at the highest indel
rates, ClustalW showed a negative relationship between FPR and
divergence---exactly opposite to the trend we expected. \prankc{}'s
FPR in the 44-taxon tree was equal to or below that of the true
alignment under almost all conditions.

The error-controlled \tpr results (Figure \ref{fig_4}C) provide a
comprehensive picture of the effect of alignment error on the detection of
\sw positive selection. The two aligners not shown in the two other
panels, MAFFT and \pranka, exhibited \tpr values intermediate to those
from ClustalW and \prankc across the range of parameters tested, with \pranka performing
better than MAFFT. As
expected, performance was very similar between aligners at very low
indel rates. At higher indel rates, most aligners yielded similar
patterns of low \tpr in the 6-taxon tree, but in the larger two trees
ClustalW and MAFFT alignments were unable to achieve high \tpr
values, presumably due largely to their elevated FPR in those trees.

It is worth noting the exceptional ability of \prankc to maintain a
very low level of false positive sites even under extremely difficult
alignment conditions. Although \prankc showed slightly elevated FPRs
at high indel rates in the 17-taxon tree, FPRs were nearly identical
to the true alignment across all simulated conditions in the 44-taxon
tree. This impressive performance suggests that, given a large enough
number of taxa, \prankc alignments would yield very few erroneous
false positives in scans for positive selection in sequences with even very
high divergence levels. Furthermore, these results showed that false
negatives contributed more than false positives to \prankc{}'s
reduction in \sw performance---a novel observation which provides
insight into the nature of \prankc alignments and their application to
\sw evolutionary analysis.

\subsection*{Effect of Alignment Filtering on \Sw Error Rates}

\begin{figure*}[t]
\begin{center}
%\includegraphics[scale=0.65]{fig5.pdf}
\includegraphics[scale=0.65]{fig5.pdf}
\end{center}
\caption{Sequences were simulated using the 17-taxon tree and a range
  of indel rates and mean path length (MPL) divergence
  levels. Alignments were inferred using (a) ClustalW or (b) \prankc,
  either left unfiltered or filtered with one of four alignment
  filters (Optimal, T-Coffee, GUIDANCE, Gblocks), and analyzed with
  SLR; true alignments were left unfiltered and separately analyzed
  with SLR. One hundred and fifty replicates were simulated for each
  set of conditions. Squares are colored according to the ratio of the
  performance of the indicated filter to the performance of the
  unfiltered ClustalW or \prankc alignment as measured by one of four
  summary statistics. In columns from left to right: the number of
  unfiltered (i.e., non-N) residues remaining in the alignment; the
  false positive rate (FPR) at the recommended cutoff threshold; the
  true positive rate (TPR) at the recommended cutoff threshold; the
  TPR at a 1\% FPR threshold (\tpr). Note that the maximum percentage
  of residues removed by filtering was capped at 50\% for all methods
  except Gblocks.}
\label{fig_5}
\end{figure*}

Having established that alignment error can lead to reduced \sw
performance through the introduction of false negatives and false
positives, we tested whether alignment filtering methods
could reduce error rates and improve the power of \sw detection of
positive selection. Using sequences simulated from the 17-taxon tree
and a range of indel rates and divergence levels, we calculated
inferred alignments using ClustalW and \prankc and applied four
filtering methods before performing the \sw analysis. Since we wished
to determine whether alignment filters either improved or worsened the
error rates and power of \sw analysis, we measured the ratio of each
performance measure to the value obtained from the equivalent
unfiltered alignments. These relative values are presented in Figure
\ref{fig_5}.

As alignment filters act through the removal of alignment residues or
columns, a certain amount of reduction in the FPR and TPR was expected
purely from the decreased amount of information available. For
example, a filter that randomly removes a fraction of residues of each
alignment would be expected to produce equal reductions in FPR and
TPR. A more effective filter may also yield a reduced TPR, but the FPR
reduction would be larger in magnitude, making the detection of
positive selection more powerful for a given error rate. Thus, a
reduced FPR is not necessarily indicative of good filtering
performance, nor is a reduced TPR necessarily indicative of poor
filtering performance. Additionally, the prevalence of false negatives
resulting from misalignment suggested that alignment filters may also
improve power by removing false negatives, perhaps by masking out
residues that were preventing positive sites from being
identified. The removal of false negatives would result in an
increased TPR, further complicating the assessment of filtering
results based on FPR or TPR alone. As a result, we focused on the
change in error-controlled \tpr{} as the best single measure of
whether a filter had successfully improved the \sw power of a dataset
since this value is sensitive to changes in both the FPR and TPR
\textcolor{red}{(note that the \tpr{} controls the FPR post-filtering,
  accounting for the tendency of filtering to reduce the FPR at a
  given cutoff threshold).}

We first examined the two controls, the unfiltered true alignment and
the inferred alignments filtered with the optimal filter (top two
rows, Figure \ref{fig_5}). The true alignment nearly always showed
smaller FPR and greater TPR and \tpr compared to the inferred
alignments, with a greater magnitude of change relative to the
ClustalW alignments than to the \prankc alignments. These scores
represented the direction and an upper limit on the magnitude of
change that might be achieved by a perfect alignment filter. One
exception to the general trend was the observation of two simulation conditions
with slightly elevated FPRs in the true alignment compared to \prankc alignments at the
lowest indel rate. This small inconsistency may be explained by
stochastic variation in false positive counts, as the absolute value
of the FPR was very low in both datasets under those conditions (on
the order of $5\times10^{-4}$, Figure \ref{fig_4}B). A similar slight FPR
elevation was also observed at the same indel rate for the optimal and
GUIDANCE filters.

Our hypothesis was that the optimal filter would show the same
direction of change in FPR and TPR as the true alignment, but with
slightly lower magnitudes. Indeed, improved \sw performance was
achieved in nearly all simulation conditions by the optimal filter,
with the magnitude of \tpr change slightly lower than for true
alignment. For ClustalW alignments the amount of improvement was quite
large, with $>$70\% increase in \tpr for nearly all conditions with an
indel rate above 0.1. The improvement was more modest for \prankc
alignments with a maximum of 15--35\% \tpr increase.

Looking at the reduction in the number of non-masked residues
remaining after filtering, we found that the optimal filter reached
the maximum of 50\% filtered residues for all ClustalW alignments with
MPL$>$1 and an indel rate$>$0.1. This meant that more than 50\% of
residues were correctly aligned across less than 50\% of the tree in
those alignments. By contrast, the optimal filter applied to \prankc
alignments only reached the maximum of 50\% filtered residues at the
highest tested divergence level and indel rate combination.

The TPR improvements achieved by the optimal filter provided some
insight into the nature of sitewise false negatives resulting from
alignment error. Two different types of alignment error might cause a
false negative at a positively-selected site: either misalignment of
one or more nonhomologous codons causing the positive signal to be
masked, or non-alignment of homologous codons causing the amount of
evolutionary information to be reduced. The former type of error would
be recoverable by alignment filtering (through removal of the codon(s)
masking the positive signal), but the latter would not. Thus, the
ability of the optimal filter to improve TPR levels across the board
provided evidence that a sizeable portion of false negatives from both
ClustalW and \prankc alignments were due to misaligned codons and thus
amenable to recovery by filtering. Although the optimal filter was
unrealistic in that it was based on perfect knowledge of which codons
were misaligned, this result provided hope that one of the other
filters might show a similar ability to recover false negative errors
from \prankc alignments.

Turning to the three filters under investigation, we found T-Coffee
and GUIDANCE both to be highly effective at improving ClustalW
alignments, with magnitudes of improvement near those of the optimal
filter. When applied to \prankc alignments, however, the two filters'
behavior diverged: T-Coffee only showed unchanged or reduced \tpr, but
GUIDANCE yielded slightly improved \tpr at high divergence levels and
indel rates, with values 5--15\% greater than the unfiltered \prankc
alignments. Both filters removed similar amounts of sequence
information and resulted in similarly reduced FPR levels, but GUIDANCE
showed a unique ability to recover false negatives from \prankc
alignments at the highest divergence levels and indel rates, and the
resulting TPR elevation appears to have been responsible for the
increased \tpr performance.

Gblocks behaved very differently from the other filters tested,
resulting in reduced FPR, TPR, and \tpr under nearly all simulation
conditions. Only at high indel rates and low divergence levels in the
ClustalW alignments did Gblocks show increased \tpr relative to the
unfiltered alignments. This poor performance was likely due to
overly-aggressive removal of alignment columns. We could not limit the
amount of sequence masked by Gblocks, so many alignments saw more than
70\% of residues removed, resulting in the loss of a large number of
correctly-aligned true positive sites. Dessimoz and Gil
\citeyearpar{Dessimoz2010Phylogenetic} found Gblocks filtering to have
a negative effetc on the accuracy of phylogenetic inference; our
results provide additional evidence in support of their finding,
suggesting that Gblocks filtering tends to reduce, rather than
increase, the power and accuracy of alignments when applied to a
number of evolutionary analyses.

There was some evidence that the column-wise nature of Gblocks
filtering was partly responsible for its poor performance in our
experiments. Since false negative errors cannot be recovered through
the removal of entire alignment columns, it made sense that the
\prankc alignments---which resulted in varying numbers of false
negatives but always very few false positives---would not see improved
performance after column-wise filtering. On the other hand, ClustalW
alignments showed a relatively constant level of false positives and
an increasing number of false negatives as divergence levels
increased. The application of a column-wise filter like Gblocks would
thus be expected to show good improvement at low divergences where
false positives dominated, but less improvement at higher divergences
where false negatives became more prominent. Indeed, this was the
pattern observed when applying Gblocks to ClustalW alignments.

Overall, our alignment filtering simulations found that filtering
methods based on GUIDANCE and T-Coffee scores had the ability to mask
away many residues causing false positives and false negatives from
ClustalW alignments, with high levels of improvement across a wide
range of divergence levels and indel rates. Filtering was less
beneficial when applied to \prankc alignments, with T-Coffee reducing
performance and GUIDANCE yielding only mild \tpr
improvements. Importantly, GUIDANCE only showed improved performance
at high divergence levels (e.g., MPL$>$1.6), well above those found in
commonly-analyzed groups of species. Thus, the use of unfiltered
\prankc alignments would yield the best performance for detecting \sw
positive selection when analyzing protein-coding sequences across a
wide range of indel rates and commonly encountered sequence divergence
levels.

\section*{Conclusions}

In this paper, we investigated the performance of sitewise detection
of positive selection under a range of tree sizes, indel rates, and
divergence levels, using simulation parameters designed to approximate
the analysis of mammalian and other typical protein-coding genes. We
evaluated the ability of four alignment methods and three alignment
filtering methods to produce alignments for detecting positively
selected sites, using the FPR, TPR, and error-controlled \tpr of the
sitewise detection of positive selection as our main measures of
performance.

The simulation results showed that alignment error can have a
measurable impact on the error rates and power of the sitewise
detection of positive selection under all but the least difficult
alignment conditions. We confirmed and extended the findings of
Fletcher and Yang \citeyearpar{Fletcher2010Effect} regarding the
relative accuracy of different aligners, showing that \prankc had the
best performance and ClustalW had the worst performance for subsequent
sitewise analysis. Notably, our simulations found that ClustalW
produced more sitewise false positives than MAFFT or \prankc even at
low divergence levels, suggesting that its use should be avoided even
when analyzing closely-related sequences. \prankc, on the other hand,
resulted in very low FPRs even at very high divergences. In
particular, when the number of sequences was large, its sitewise FPRs
were virtually indistinguishable from the true alignment's.

The very low FPRs observed for \prankc alignments conflicted somewhat
with the results of Fletcher and Yang
\citeyearpar{Fletcher2010Effect}, who found that the FPRs for the
branch-site test were not under control even with \prankc
alignments. This apparent discrepancy can be explained by different
sensitivities to alignment error: the branch-site test would yield
false positives when misalignment causes apparent positive selection
along only the foreground branch, while the SLR sitewise test would produce
false positives only when misalignment causes a signal of positive
selection strong enough to overpower the non-positive signal
throughout the tree. This effect stems from the different biological
hypotheses tested by the two methods; their differential sensitivity
to misalignment underscores the necessity of considering the
biological sensitivity and robustness to alignment error when applying
either of these tests to detect positive selection within an
alignment.

Despite producing very low FPRs, \prankc alignments still resulted in
an increased number of false negatives compared to the true
alignment. We showed that some of these false negatives were possible
to recover as true positives through alignment filtering, and we found
that both the optimal filter and GUIDANCE were able to successfully
recover these false negatives at high divergence levels, resulting in
small but measurable performance improvements over the unfiltered
\prankc alignments.

Among the aligners and filters tested here, we can recommend the use
of unfiltered \prankc alignments for the detection of sitewise
positive selection in all but the most divergent phylogenies. GUIDANCE
showed some ability to improve sitewise power under difficult
alignment conditions at high divergence levels, but this improvement
was achieved largely through the recovery of false negatives as
opposed to the elimination of false positives. For an analysis where
the control of false positives is paramount, the added computational
expense of running many bootstrap alignment replicates (as is done
GUIDANCE) may not be worth the possibility of a slight increase in
power. The development of efficient and effective alignment filters,
particularly with an eye towards recovering false negatives as
identified in this study, will be an interesting area for future
research.

Our simulations did not include fully biologically realistic models of
spatial or temporal variability in the rate of indel formation or in
the distribution of selective pressures \citep{Whelan2008Spatial}, but
we do not expect that such heterogeneity would affect our main
conclusions regarding the relative performance of different aligners
and filters. The trends we observed were consistent across a wide
range of parameter values, suggesting that they reflect fundamental
differences in each method's ability to align or filter sequences as
opposed to artifacts due to the relative simplicity of our
simulations.

However, the apropriateness of our simulation scheme should be
critically considered when evaluating our specific power and error
rate estimates in the context of real-world data analysis. Functional
differences between genes may influence the $\omg$ distribution and
indel rate---with some genes or domains showing lower or higher
tolerance to indels and fewer or more neutrally-evolving sites than
modeled in our simulations---making false positive results more or
less likely. Second, species-level differences may also be important,
as the efficacy of natural selection is highly dependent on effective
population size \citep{Ellegren2009Selection}. For example, proteins
evolving in \Dr species with a high effective population size should
experience stronger positive and purifying selection than in mammals,
potentially leading to increased power and reduced error rates when
compared to our simulations based on a mammalian-like $\omg$
distribution.

We showed here that even relatively simple evolutionary simulation
experiments could sensitively assess the performance characteristics
of different aligners, provide quantitative insight into the practical
effects of alignment error, and suggest areas for future development
of alignment and filtering methods. In the future, we expect the
development of more realistic simulations for protein
evolution---perhaps incorporating structurally-motivated and
empirically validated models of mutation, indel formation and
constraint---to further increase the applicability and accuracy of
such experiments, \textcolor{red}{and we believe that flexible and
  accessible simulation programs such as INDELible
  \citep{Fletcher2009INDELible} and PhyloSim \citep{Sipos2011PhyloSim}
  will play an important role in the quantitative assessment of
  alignment algorithms and alignment-dependent comaprative analyses.}

As genomes rapidly accumulate in the databases and
large-scale analyses become the norm, we hope that the development and
application of alignment methods, which are arguably the most
important step in any evolutionary analysis, will be based on a
rigorous understanding of their behavior and performance when applied
to a wide variety of evolutionary analyses.

\section*{Acknowledgments}

GJ was funded by a Gates Cambridge Trust Scholarship and is a member
of Darwin College, University of Cambridge.

\bibliographystyle{mbe3} \fontsize{9}{10}\selectfont% \itemsep 0pt

\bibliography{slrsim}
\end{document}