### 2011-07-21 projects/slrsim/paper/supp_slrsim.tex

\documentclass{article}
\usepackage{natbib,graphicx,dblfloatfix,multirow,array,threeparttable,xspace}
\usepackage[landscape,hmargin=1.5cm,vmargin=1.5cm]{geometry}

\newcommand{\pranka}{PRANK$_{\textrm{AA}}$\xspace}

\begin{document}

\makeatletter 
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\thefigure}{S\@arabic\c@figure}

\begin{figure}[t]
\begin{center}
%\includegraphics[scale=0.65]{supp_fig1.pdf}
\includegraphics[scale=0.8]{supp_fig1.pdf}
\end{center}
\caption{This figure depicts the same simulations and uses the same
  formatting as Figure 4, except that results for MAFFT and \pranka
  have been added to sections (a) and (b) and results using ClustalW +
  GUIDANCE have been added to all sections.}
\label{supp_fig1}
\end{figure}

\end{document}